A Structural and Functional Bioinformatics Study of QTY-designed Retinylidene Proteins

### 2025-05-16 Paper/Manuscript/manuscript.tex
Submit to QRB.

Retinylidene proteins are retinal-binding light-sensitive proteins found in organisms ranging from microbes to human. Microbial opsins have been utilized in optogenetics, while animal opsins are essential for vision and light-dependent metabolic functions. However, retinylidene proteins have hydrophobic transmembrane (TM) domains, which makes them challenging to study. In this structural and functional bioinformatics study, we used the QTY (Glutamine, Threonine, Tyrosine) code to design water-soluble QTY analogs of retinylidene proteins, including 9 human and 3 microbial opsins. We provide superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water-soluble QTY analogs, and experimentally-determined structures when available. We also provide a comparison of surface hydrophobicity of the variants. Despite significant changes to the protein sequence (35.53\% to 50.24\% in the TM domain), protein characteristics and structures are well-preserved. Furthermore, we run molecular dynamics (MD) simulations of native and QTY-designed OPN2 (rhodopsin) and analyze their response to the isomerization of 11-cis-retinal to all-trans-retinal. The results show that the QTY analog has similar functional behavior to the native protein. The findings of this study indicate that the QTY code can be used as a robust tool to design water-solubility retinylidene proteins. These have potential applications in protein studies, therapeutic treatments, and bioengineering.

@@

Retinylidene proteins are photochemically reactive proteins that are bound to or can bind to retinal (vitamin A aldehyde) as their chromophore \citep{Spudich_2000}. We will use the term ``opsin'' interchangeably with ``retinylidene protein'', despite the fact that ``opsin'' sometimes refers specifically to the chromophore-free apoprotein. Retinylidene proteins are divided into two groups, animal opsins and microbial opsins, which are evolutionarily distinct but share common characteristics such as having 7 transmembrane (TM) domains and a retinal-binding lysine residue \citep{Yee_2013, Spudich_2000}. In this study, we will consider 9 animal opsins and 3 microbial opsins.

@@

In this study, we selected 9 opsins expressed in the human nervous system: OPN1MW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04001/entry}{P04001}), OPN1LW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04000/entry}{P04001}), OPN1SW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P03999/entry}{P03999}), OPN2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P08100/entry}{P04001}), OPN3 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9H1Y3/entry}{Q9H1Y3}), OPN4 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9UHM6/entry}{Q9UHM6}), OPN5 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q6U736/entry}{Q6U736}), RGR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P47804/entry}{P47804}) and RRH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/O14718/entry}{O14718}). They belong to several evolutionarily distinct families of animal opsins \citep{Terakita_2005}.

@@

\textbf{OPN2} (Opsin 2), also known as rhodopsin, is expressed in retinal rod photoreceptors and is responsible for vision at low light intensity \citep{Hubbard_1958}. Certain variants lead to autosomal recessive or autosomal dominant retinitis pigmentosa and congenital stationary night blindness \citep{Fanelli_2021}. OPN2 is a representative animal opsin, one of the first studied. In fact, bovine OPN2 is the first opsin to be sequenced \citep{Nathans_1984}, as well as the first GPCR whose crystal structure was resolved experimentally \citep{Palczeski_2000}. Many studies on the functional mechanisms of animal opsins also focus on OPN2. Consequently, we choose OPN2 to conduct a functional analysis in this study, in order to further explore the effectiveness of the QTY code in redesigning retinylidene proteins.

@@

In this study, we selected 3 microbial opsins: BACR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P02945/entry}{P02945}), BACH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/B0R2U4/entry}{B0R2U4}), ChR2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q8RUT8/entry}{Q8RUT8}).

@@

We have asked whether retinylidene proteins could be redesigned to be more soluble. Retinylidene proteins are all integral membrane proteins with seven transmembrane alpha helices embedded in a lipid bilayer. Because of the hydrophobic properties of the transmembrane domains, they are not water-soluble without the aid of detergents.

@@

Instead of taking a computational approach, we applied the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) and glutamine (Q); isoleucine (I) / valine (V) and threonine (T); and phenylalanine (F) and tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. Zhang et al. initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. They have also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. In addition to GPCR, they have also used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor has been successfully used in biomimetic sensors \citep{Qing_2023}. They have also applied the rQTY (reverse QTY) code to design self-assembled nanoparticles \citep{Meng_2023}.

@@

In this paper, we apply the QTY code to redesign 9 human opsins and 3 microbial opsins. We provide the superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water-soluble QTY variants, and experimentally-determined structures when available. We also provide a comparison of the surface hydrophobicity of the variants. Furthermore, we run MD simulations of native and QTY-designed OPN2 and analyze their response to the 11-cis to all-trans isomerization of the retinal chromophore.

@@

We predicted the structures of native proteins and their QTY variants using the AlphaFold3 website (\url{https://alphafoldserver.com}) \citep{Abramson_2024}. For microbial opsins, the dimers and trimers had identical subunits, i.e., they were homodimers and homotrimers. In AlphaFold3, this was achieved by altering the number of protein copies.

@@

PDB files for native protein structures determined experimentally by X-ray diffraction or electron microscopy were taken from RCSB PDB: OPN2 (PDB ID: \href{https://www.rcsb.org/structure/5W0P}{5W0P}), BACR (PDB ID: \href{https://www.rcsb.org/structure/7XJC}{7XJC}), BACH (PDB ID: \href{https://www.rcsb.org/structure/2JAF}{2JAF}), ChR2 (PDB ID: \href{https://www.rcsb.org/structure/8ZAN}{8ZAN}). The AlphaFold3-predicted native and QTY variants were taken directly from the most probable predicted structure. Superpositions were performed via the command ``super'' in PyMOL (\url{https://pymol.org}). We remove unstructured loops at the N and C terminals for the sake of clarity.

@@

We used PyMOL (\url{https://pymol.org}) to superpose the native predicted protein structures, their QTY variants, and the experimentally determined structures for the proteins where these existed. We then used UCSF ChimeraX (\url{https://www.cgl.ucsf.edu/chimerax/}) to render each protein model with hydrophobicity patches.

@@

Regarding QTY-designed OPN2, we found that the helix 9 of the QTY analog was flung up and adhered to the originally transmembrane domain, which disrupted the normal conformation. Therefore, this helix was truncated. The system was constructed directly using GROMACS. The protein was centered in a rectangular box with dimensions equal to those of native OPN2. The equilibration involved a sequence of simulations: one 250-ps NVT equilibration, followed by one 125-ps and one 1500-ps NPT equilibrations, totaling an equilibration duration that matches that of the natural OPN2. All other configurations and parameters are identical to those of the original OPN2.

After the simulations, we extracted frames from the trajectory to inspect the structures before and after retinal isomerization. We also calculated the RMSD (``gmx rms'' command), number of hydrogen bonds (``gmx hbond'' command) and water molecules (``gmx select'' command) inside the binding pocket, interaction energy (including short-range Coulombic interaction energy and short-range Lennard-Jones energy) (``gmx energy'' command), the RMSF (``gmx rmsf'' command), and radius of gyration (``gmx gyrate'' command). The detailed commands are publicly available in the online database for this study.

@@

The transmembrane segments of membrane proteins are hydrophobic in nature. This leads to a challenge to study their structure and function. Traditionally, detergent must be used to purify them. We consider a different approach, i.e., via systematic soluble protein design. There exists structural similarities, as can be observed on high-resolution electron density maps, between certain hydrophobic and hydrophilic amino acids: leucine (L) and glutamine (Q), isoleucine (I) or valine (V) and threonine (T), and phenylalanine (F) with tyrosine (Y). This fact enables systematic replacement of L with Q, I/V with T, and F with Y in all transmembrane segments of the membrane protein, which we named the QTY code. QTY analogs are less hydrophobic than native membrane proteins. Although their amino acid sequences have been significantly changed, they still exhibit relatively preserved structure, isoelectric points (pI), and molecular weights (MW) compared to the native membrane proteins (Table~\ref{tb:characteristics}).

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. We used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superpose very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, we also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) are small, from 0.307{\AA} to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999\AA). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

All 9 of these human opsins belong to the same large family, i.e. animal opsins. However, they belong to different subfamilies and have structural differences from each other. In order to investigate the extent to which the water-soluble QTY analogs preserve these differences, we also conducted pairwise superpositions within all 9 native human opsins and within all 9 QTY analogs of human opsins (Figure~\ref{fig:humansup} \textbf{j)} and \textbf{k)}). As observed in the tables, there is a large degree of similarity between the RMSD between a pair of native proteins and that between the corresponding pair of QTY analogs. Therefore, we conclude that the structural differences are relatively well-preserved before and after the QTY substitution. However, whether the functional differences are preserved remains a problem to study.

@@

We used AlphaFold3 to predict the structure of monomers of the 3 native microbial opsins, their water-soluble QTY analogs. Since microbial opsins often have to form oligomers to be functional, we also predicted the structure of native and QTY variants of BACR trimers, BACH trimers, and ChR2 dimers. These predicted structures are also superposed with experimentally determined structures (X-ray diffraction or electron microscopy). The structures superpose very well (Figure~\ref{fig:microbialsup}). Overall, the root mean square differences (RMSD) are small, with the highest value being 0.685\AA. This shows that the water-soluble QTY analogs are likely capable of forming oligomers with structures similar to those of native proteins.

@@

We investigate the surface hydrophobicity of water-soluble QTY-analogs in comparison with the native proteins. After applying the QTY code to respectively replace the hydrophobic amino acids L, I/V, and F with hydrophilic amino acids glutamine (Q), threonine (T), and tyrosine (Y), the hydrophobic surface areas are significantly reduced. More importantly, since the electron density maps of the corresponding amino acids are similar, the alpha helices in QTY analogs have a similar contour to those in the native proteins. The QTY analogs have successfully retained its structural integrity and stability.

@@

However, AlphaFold3 still has limitations that have been encountered in this study. Although AlphaFold3 allows the structural prediction of complexes that include small-molecule ligands, it only supports a small range of ligands, which does not include retinal. We were therefore unable to predict the structure of chromophore-bound states of retinylidene proteins with AlphaFold3. Consequently, we chose another approach, namely molecular dynamics simulation, in order to investigate the behavior of QTY opsin analogs when they are bound to the chromophore.

@@

QTY substitutions do not introduce changes to the essential NPxxY motif or the retinal-binding lysine of human retinylidene proteins. Therefore, we propose that QTY analogs of human opsins may have conserved functions and carry out a molecular dynamics (MD) simulation. We record and analyze the behavior of native OPN2 and its QTY analog before and after the 11-cis to all-trans isomerization.

@@

Zooming in on the retinal-binding pocket, we observe that retinal is held within its binding pocket by several forces in both the native and QTY analog: the counterions of the retinal Schiff base (RSB) (113E and 181E), the hydrophobic interactions with neighboring residues (for example, 265W, 268Y, and the 187--189 beta strand), and, sometimes, steric collisions (Figure~\ref{fig:binding}). Since retinal is a ligand that binds in an internal pocket, we were unable to calculate the Gibbs free energy of its binding. Nevertheless, the distribution of interaction energies are found to be similar in native OPN2 and the water-soluble QTY analog, except for small differences: native OPN2 experiences a change in interaction energy at about $t=50$ns due to the entrance of water near 113E; the QTY analog has more negative interaction energy at residues 208 and 212 due to phenylalanine (F) to tyrosine (Y) substitutions. Besides from these, the similarity is a strong suggestion of functional conservation. Finally, we observe very different hydrogen bonding and water molecule patterns inside the binding pocket, which is mainly due to several F to Y substitutions in that region. This may cause retinal to interact with adjacent residues in a subtly different way, and may lead to a different absorption peak since the environment of the RSB, i.e., the binding pocket, determines the spectral characteristic of opsins \citep{Fenno_2011}.

Overall, we observe that QTY-designed OPN2 has relatively similar behavior to the native variant despite certain small differences caused by decreased hydrophobicity. The function of OPN2$^{\textrm{QTY}}$ awaits to be verified by experiments.

@@

In this study, we selected 12 retinylidene proteins, including 9 human opsins (OPN1MW, OPN1LW, OPN1SW, OPN2, OPN3, OPN4, OPN5, RGR, and RRH) and 3 microbial opsins (BACR, BACH, and ChR2), that have clinical implications and various applications. We applied the QTY code to convert the hydrophobic transmembrane alpha-helices into hydrophilic ones and thus create water-soluble QTY analogs of the proteins. Then, we used AlphaFold3 to predict the structures of the native proteins as well as the QTY analogs, superposed them, and found that the structure of the proteins are well conserved despite substantial residue substitutions. We also inspected the surface hydrophobicity of the proteins and found a great decrease in hydrophobic patches in the transmembrane surface. Next, we chose a representative retinylidene protein, OPN2, and employed molecular dynamics simulations to investigate the behavior of native and QTY analog of OPN2. We found that the QTY analog is capable of conformational change and effective protein-ligand interaction. These findings reveal that QTY-designed OPN2, and retinylidene proteins in general, may have conserved structure and function. We believe that these water-soluble QTY variants may have potentials in research, therapeutic treatments, and bioengineering.

>>>>

Retinylidene proteins are retinal-binding light-sensitive proteins found in organisms ranging from microbes to human. Microbial opsins have been utilized in optogenetics, while animal opsins are essential for vision and light-dependent metabolic functions. However, retinylidene proteins have hydrophobic transmembrane (TM) domains, which makes them challenging to study. In this structural and functional bioinformatics study, I used the QTY (Glutamine, Threonine, Tyrosine) code to design water-soluble QTY analogs of retinylidene proteins, including 9 human and 3 microbial opsins. I provide superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water-soluble QTY analogs, and experimentally-determined structures when available. I also provide a comparison of surface hydrophobicity of the variants. Despite significant changes to the protein sequence (35.53\% to 50.24\% in the TM domain), protein characteristics and structures are well-preserved. Furthermore, I run molecular dynamics (MD) simulations of native and QTY-designed OPN2 (rhodopsin) and analyze their response to the isomerization of 11-cis-retinal to all-trans-retinal. The results show that the QTY analog has similar functional behavior to the native protein. The findings of this study indicate that the QTY code can be used as a robust tool to design water-solubility retinylidene proteins. These have potential applications in protein studies, therapeutic treatments, and bioengineering.

@@

Retinylidene proteins are photochemically reactive proteins that are bound to or can bind to retinal (vitamin A aldehyde) as their chromophore \citep{Spudich_2000}. I will use the term ``opsin'' interchangeably with ``retinylidene protein'', despite the fact that ``opsin'' sometimes refers specifically to the chromophore-free apoprotein. Retinylidene proteins are divided into two groups, animal opsins and microbial opsins, which are evolutionarily distinct but share common characteristics such as having 7 transmembrane (TM) domains and a retinal-binding lysine residue \citep{Yee_2013, Spudich_2000}. In this study, I will consider 9 animal opsins and 3 microbial opsins.

@@

In this study, I selected 9 opsins expressed in the human nervous system: OPN1MW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04001/entry}{P04001}), OPN1LW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04000/entry}{P04001}), OPN1SW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P03999/entry}{P03999}), OPN2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P08100/entry}{P04001}), OPN3 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9H1Y3/entry}{Q9H1Y3}), OPN4 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9UHM6/entry}{Q9UHM6}), OPN5 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q6U736/entry}{Q6U736}), RGR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P47804/entry}{P47804}) and RRH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/O14718/entry}{O14718}). They belong to several evolutionarily distinct families of animal opsins \citep{Terakita_2005}.

@@

\textbf{OPN2} (Opsin 2), also known as rhodopsin, is expressed in retinal rod photoreceptors and is responsible for vision at low light intensity \citep{Hubbard_1958}. Certain variants lead to autosomal recessive or autosomal dominant retinitis pigmentosa and congenital stationary night blindness \citep{Fanelli_2021}. OPN2 is a representative animal opsin, one of the first studied. In fact, bovine OPN2 is the first opsin to be sequenced \citep{Nathans_1984}, as well as the first GPCR whose crystal structure was resolved experimentally \citep{Palczeski_2000}. Many studies on the functional mechanisms of animal opsins also focus on OPN2. Consequently, I choose OPN2 to conduct a functional analysis in this study, in order to further explore the effectiveness of the QTY code in redesigning retinylidene proteins.

@@

In this study, I selected 3 microbial opsins: BACR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P02945/entry}{P02945}), BACH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/B0R2U4/entry}{B0R2U4}), ChR2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q8RUT8/entry}{Q8RUT8}).

@@

I have asked whether retinylidene proteins could be redesigned to be more soluble. Retinylidene proteins are all integral membrane proteins with seven transmembrane alpha helices embedded in a lipid bilayer. Because of the hydrophobic properties of the transmembrane domains, they are not water-soluble without the aid of detergents.

@@

Instead of taking a computational approach, I applied the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) and glutamine (Q); isoleucine (I) / valine (V) and threonine (T); and phenylalanine (F) and tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. Zhang et al. initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. Zhang's team have also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. In addition to GPCR, they have also used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor has been successfully used in biomimetic sensors \citep{Qing_2023}. They have also applied the rQTY (reverse QTY) code to design self-assembled nanoparticles \citep{Meng_2023}.

@@

In this paper, I apply the QTY code to redesign 9 human opsins and 3 microbial opsins. I provide the superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water-soluble QTY variants, and experimentally-determined structures when available. I also provide a comparison of the surface hydrophobicity of the variants. Furthermore, I run MD simulations of native and QTY-designed OPN2 and analyze their response to the 11-cis to all-trans isomerization of the retinal chromophore.

@@

I predicted the structures of native proteins and their QTY variants using the AlphaFold3 website (\url{https://alphafoldserver.com}) \citep{Abramson_2024}. For microbial opsins, the dimers and trimers had identical subunits, i.e., they were homodimers and homotrimers. In AlphaFold3, this was achieved by altering the number of protein copies.

@@

PDB files for native protein structures determined experimentally by X-ray diffraction or electron microscopy were taken from RCSB PDB: OPN2 (PDB ID: \href{https://www.rcsb.org/structure/5W0P}{5W0P}), BACR (PDB ID: \href{https://www.rcsb.org/structure/7XJC}{7XJC}), BACH (PDB ID: \href{https://www.rcsb.org/structure/2JAF}{2JAF}), ChR2 (PDB ID: \href{https://www.rcsb.org/structure/8ZAN}{8ZAN}). The AlphaFold3-predicted native and QTY variants were taken directly from the most probable predicted structure. Superpositions were performed via the command ``super'' in PyMOL (\url{https://pymol.org}). I remove unstructured loops at the N and C terminals for the sake of clarity.

@@

I used PyMOL (\url{https://pymol.org}) to superpose the native predicted protein structures, their QTY variants, and the experimentally determined structures for the proteins where these existed. I then used UCSF ChimeraX (\url{https://www.cgl.ucsf.edu/chimerax/}) to render each protein model with hydrophobicity patches.

@@

Regarding QTY-designed OPN2, I found that the helix 9 of the QTY analog was flung up and adhered to the originally transmembrane domain, which disrupted the normal conformation. Therefore, this helix was truncated. The system was constructed directly using GROMACS. The protein was centered in a rectangular box with dimensions equal to those of native OPN2. The equilibration involved a sequence of simulations: one 250-ps NVT equilibration, followed by one 125-ps and one 1500-ps NPT equilibrations, totaling an equilibration duration that matches that of the natural OPN2. All other configurations and parameters are identical to those of the original OPN2.

After the simulations, I extracted frames from the trajectory to inspect the structures before and after retinal isomerization. I also calculated the RMSD (``gmx rms'' command), number of hydrogen bonds (``gmx hbond'' command) and water molecules (``gmx select'' command) inside the binding pocket, interaction energy (including short-range Coulombic interaction energy and short-range Lennard-Jones energy) (``gmx energy'' command), the RMSF (``gmx rmsf'' command), and radius of gyration (``gmx gyrate'' command). The detailed commands are publicly available in the online database for this study.

@@

The transmembrane segments of membrane proteins are hydrophobic in nature. This leads to a challenge to study their structure and function. Traditionally, detergent must be used to purify them. Zhang et al. considered a different approach, i.e., via systematic soluble protein design. There exists structural similarities, as can be observed on high-resolution electron density maps, between certain hydrophobic and hydrophilic amino acids: leucine (L) and glutamine (Q), isoleucine (I) or valine (V) and threonine (T), and phenylalanine (F) with tyrosine (Y). This fact enables systematic replacement of L with Q, I/V with T, and F with Y in all transmembrane segments of the membrane protein, which Zhang et al. named the QTY code. QTY analogs are less hydrophobic than native membrane proteins. Although their amino acid sequences have been significantly changed, they still exhibit relatively preserved structure, isoelectric points (pI), and molecular weights (MW) compared to the native membrane proteins (Table~\ref{tb:characteristics}).

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. I used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superpose very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, I also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) are small, from 0.307{\AA} to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999\AA). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

All 9 of these human opsins belong to the same large family, i.e. animal opsins. However, they belong to different subfamilies and have structural differences from each other. In order to investigate the extent to which the water-soluble QTY analogs preserve these differences, I also conducted pairwise superpositions within all 9 native human opsins and within all 9 QTY analogs of human opsins (Figure~\ref{fig:humansup} \textbf{j)} and \textbf{k)}). As observed in the tables, there is a large degree of similarity between the RMSD between a pair of native proteins and that between the corresponding pair of QTY analogs. Therefore, I conclude that the structural differences are relatively well-preserved before and after the QTY substitution. However, whether the functional differences are preserved remains a problem to study.

@@

I used AlphaFold3 to predict the structure of monomers of the 3 native microbial opsins, their water-soluble QTY analogs. Since microbial opsins often have to form oligomers to be functional, I also predicted the structure of native and QTY variants of BACR trimers, BACH trimers, and ChR2 dimers. These predicted structures are also superposed with experimentally determined structures (X-ray diffraction or electron microscopy). The structures superpose very well (Figure~\ref{fig:microbialsup}). Overall, the root mean square differences (RMSD) are small, with the highest value being 0.685\AA. This shows that the water-soluble QTY analogs are likely capable of forming oligomers with structures similar to those of native proteins.

@@

I investigate the surface hydrophobicity of water-soluble QTY-analogs in comparison with the native proteins. After applying the QTY code to respectively replace the hydrophobic amino acids L, I/V, and F with hydrophilic amino acids glutamine (Q), threonine (T), and tyrosine (Y), the hydrophobic surface areas are significantly reduced. More importantly, since the electron density maps of the corresponding amino acids are similar, the alpha helices in QTY analogs have a similar contour to those in the native proteins. The QTY analogs have successfully retained its structural integrity and stability.

@@

However, AlphaFold3 still has limitations that have been encountered in this study. Although AlphaFold3 allows the structural prediction of complexes that include small-molecule ligands, it only supports a small range of ligands, which does not include retinal. I were therefore unable to predict the structure of chromophore-bound states of retinylidene proteins with AlphaFold3. Consequently, I chose another approach, namely molecular dynamics simulation, in order to investigate the behavior of QTY opsin analogs when they are bound to the chromophore.

@@

QTY substitutions do not introduce changes to the essential NPxxY motif or the retinal-binding lysine of human retinylidene proteins. Therefore, I propose that QTY analogs of human opsins may have conserved functions and carry out a molecular dynamics (MD) simulation. I record and analyze the behavior of native OPN2 and its QTY analog before and after the 11-cis to all-trans isomerization.

@@

Zooming in on the retinal-binding pocket, I observe that retinal is held within its binding pocket by several forces in both the native and QTY analog: the counterions of the retinal Schiff base (RSB) (113E and 181E), the hydrophobic interactions with neighboring residues (for example, 265W, 268Y, and the 187--189 beta strand), and, sometimes, steric collisions (Figure~\ref{fig:binding}). Since retinal is a ligand that binds in an internal pocket, I were unable to calculate the Gibbs free energy of its binding. Nevertheless, the distribution of interaction energies are found to be similar in native OPN2 and the water-soluble QTY analog, except for small differences: native OPN2 experiences a change in interaction energy at about $t=50$ns due to the entrance of water near 113E; the QTY analog has more negative interaction energy at residues 208 and 212 due to phenylalanine (F) to tyrosine (Y) substitutions. Besides from these, the similarity is a strong suggestion of functional conservation. Finally, I observe very different hydrogen bonding and water molecule patterns inside the binding pocket, which is mainly due to several F to Y substitutions in that region. This may cause retinal to interact with adjacent residues in a subtly different way, and may lead to a different absorption peak since the environment of the RSB, i.e., the binding pocket, determines the spectral characteristic of opsins \citep{Fenno_2011}.

Overall, I observe that QTY-designed OPN2 has relatively similar behavior to the native variant despite certain small differences caused by decreased hydrophobicity. The function of OPN2$^{\textrm{QTY}}$ awaits to be verified by experiments.

@@

In this study, I selected 12 retinylidene proteins, including 9 human opsins (OPN1MW, OPN1LW, OPN1SW, OPN2, OPN3, OPN4, OPN5, RGR, and RRH) and 3 microbial opsins (BACR, BACH, and ChR2), that have clinical implications and various applications. I applied the QTY code to convert the hydrophobic transmembrane alpha-helices into hydrophilic ones and thus create water-soluble QTY analogs of the proteins. Then, I used AlphaFold3 to predict the structures of the native proteins as well as the QTY analogs, superposed them, and found that the structure of the proteins are well conserved despite substantial residue substitutions. I also inspected the surface hydrophobicity of the proteins and found a great decrease in hydrophobic patches in the transmembrane surface. Next, I chose a representative retinylidene protein, OPN2, and employed molecular dynamics simulations to investigate the behavior of native and QTY analog of OPN2. I found that the QTY analog is capable of conformational change and effective protein-ligand interaction. These findings reveal that QTY-designed OPN2, and retinylidene proteins in general, may have conserved structure and function. I believe that these water-soluble QTY variants may have potentials in research, therapeutic treatments, and bioengineering.
